---
title: "What makes a country happy?"
author: "Stella Lee, Tanvi Jindal"
output:
  ioslides_presentation: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Research Questions
We want to study how a country's happiness can be influenced by various measures of a country's attributes.

- How does a country’s happiness index get influenced by the following factors:
    - GDP per capita
    - Life Expectancy
    - Government Corruption
    - Region
- How does time affect this?

## Source
To study our questions, we used data from different sources, as numerated below:

- Happiness data: Happy Planet Index
- GDP per Capita (PPP): Happy Planet Index
- Life Expectancy: World Happiness Report
- Corruption: World Happiness Report

We used data from 2009, 2012 and 2016 because those were the years for which Happy Planet Index data were available. Moreover, the three time periods show a general trends taking place over a decade.
  
## Map Plotting Happiness, GDP per capita, Life Expectancy, and Corruption for 2009

```{r, warning = FALSE, message = FALSE}
library(plotly)
library(ggplot2)
library(readr)
library(dplyr)
library(tidyr)
library(maps) 
#install.packages("scatterplot3d")
library(scatterplot3d)

happy09 <- read.csv("https://raw.githubusercontent.com/stellasylee/WorldHappinessData/master/Data/2009HPI.csv")
happy12 <- read.csv("https://raw.githubusercontent.com/stellasylee/WorldHappinessData/master/Data/2012HPI.csv")
happy16 <- read.csv("https://raw.githubusercontent.com/stellasylee/WorldHappinessData/master/Data/2016HPI.csv")
variables <- read.csv("https://raw.githubusercontent.com/stellasylee/WorldHappinessData/master/Data/OtherVariables.csv")
# 2009
happy09 <- select(happy09, country, HPI, GDP.per.capita....PPP.)
names(happy09) <- c("country", "HPI09", "GDP09")
happy09$country <- as.character(happy09$country)
happy09$country[happy09$country == "Burma"] <- "Myanmar"
happy09$country[happy09$country == "Korea"] <- "South Korea"
happy09$country[happy09$country == "Congo"] <- "Republic of Congo" 
happy09$country[happy09$country == "Congo, Dem. Rep. of the"] <- "Democratic Republic of the Congo"

# 2012
happy12 <- select(happy12, Country, Happy.Planet.Index, GDP.capita...PPP.)
names(happy12) <- c("country", "HPI12", "GDP12")
happy12$country <- as.character(happy12$country)
happy12$country[happy12$country == "Korea"] <- "South Korea"
happy12$country[happy12$country == "Congo"] <- "Republic of Congo" 
happy12$country[happy12$country == "Congo, Dem. Rep. of the"] <- "Democratic Republic of the Congo"

# 2016
happy16 <- select(happy16, Country, Region, Happy.Planet.Index, X.GDP.capita...PPP.. )
names(happy16) <- c("country","region","HPI16", "GDP16")
happy16$country <- as.character(happy16$country)

# variables
variables <- filter(variables, year == "2009" | year == "2012" | year == "2016")
var <- select (variables, country, year, Healthy.life.expectancy.at.birth, Perceptions.of.corruption)
names(var) <- c("country", "year", "lifeExpect", "corruption") 
var$year <- as.character(var$year)
var$country <- as.character(var$country)
var$country[var$country == "Congo (Brazzaville)"] <- "Republic of Congo" 
var$country[var$country == "Congo (Kinshasa)"] <- "Democratic Republic of the Congo"
var$country[var$country == "Hong Kong S.A.R. of China"] <- "Hong Kong"
var$country[var$country == "United States"] <- "United States of America"
var$country[var$country == "Palestinian Territories"] <- "Palestine"

# Combine
happy <- full_join(x=happy16, y=happy12, by = "country")
happy <- full_join(x=happy, y=happy09, by = "country")
happy$country[happy09$country == "Burma"] <- "Myanmar"
happy$GDP09 <- parse_number(happy$GDP09)
happy$GDP12 <- parse_number(happy$GDP12)
happy$GDP16 <- parse_number(happy$GDP16)

happy <- gather(data = happy, key = year, value = Index, 3:8)
happy <- separate(happy, year, into = c("variable", "year"), sep = -3)
happy$year<- paste0("20", happy$year)
happy <- spread (data = happy, key = variable, value = Index)
happy <- left_join(x = happy, y = var, by = c("country", "year"))

World <- map_data("world")
# Change the column and row for matching with the Happiness dataset
names(World) <- c("long", "lat", "group", "order", "country", "subregion")
World$country[World$country == "USA"] <- "United States of America" 
World$country[World$country == "UK"] <- "United Kingdom" 
World$country[World$subregion == "Hong Kong"] <- "Hong Kong" 
World$country[World$country == "Trinidad" | World$country == "Tobago" ] <- "Trinidad and Tobago" 

# Join the map, happiness data for visualization
final09 <- filter(happy, year == "2009")
map09 <- full_join(World, final09, by = "country")
GDPMap09 <- ggplot() + geom_polygon(data=map09, 
          aes(x=long, y=lat, group=group, fill = log(GDP)), color="white", size = 0.2) +
  scale_fill_gradientn(colours = rainbow(7))+guides(fill=FALSE)

lifeMap09 <- ggplot() + geom_polygon(data=map09, 
          aes(x=long, y=lat, group=group, fill = lifeExpect), 
          color="white", size = 0.2) + 
  scale_fill_gradientn(colours = rainbow(7))+guides(fill=FALSE)

corrupMap09 <- ggplot() + geom_polygon(data=map09, 
          aes(x=long, y=lat, group=group, fill = corruption), 
          color="white", size = 0.2) + 
  scale_fill_gradientn( colours = rainbow(7)) +guides(fill=FALSE)

happyMap09 <- ggplot() + geom_polygon(data=map09, 
          aes(x=long, y=lat, group=group, fill = HPI), 
          color="white", size = 0.2) + 
  scale_fill_gradientn(colours = rainbow(7))+guides(fill=FALSE)

subplot(
  ggplotly(GDPMap09),
  ggplotly(happyMap09),
  ggplotly(lifeMap09),
  ggplotly(corrupMap09),
  nrows = 2, shareX = TRUE, shareY = TRUE, titleY = FALSE, titleX = FALSE
) %>% layout(title = "Happy Planet Index 2009 ", annotations = list(
  list(yanchor = "middle", xanchor = "left", align = "center",x= 0.15, y = 0.54, text = "GDP per capita",showarrow=FALSE,xref = "paper", yref="paper"),
  list(yanchor = "middle", xanchor = "right", align = "center",x= 0.75, y = 0.54, text = "HPI",showarrow=FALSE,xref = "paper", yref="paper"),
  list(yanchor = "middle", xanchor = "left", align = "center",x=0.15, y = 0.0, text = "Life Expectancy",showarrow=FALSE,xref = "paper", yref="paper"),
  list(yanchor = "middle", xanchor = "right", align = "center",x=0.8, y = 0.0, text = "Corruption",showarrow=FALSE,xref = "paper", yref="paper")
))
```

## Interactive Map
```{r}

```

## Scatterplot of Happiness vs GDP 
```{r}
ggplot(data=final09)+  geom_point(aes(x=log(GDP), y=HPI, colour = region))
```
## Scatterplot of Happiness vs Life Expectancy
```{r}
ggplot(data=final09)+  geom_point(aes(x=lifeExpect, y=HPI, colour = region))
```

## 3D Scatterplot
```{r}
region <- as.numeric(final09$region)
scatterplot3d(log(final09$GDP), final09$lifeExpect, final09$HPI, color = region, type = "h", main="3D Scatter Plot", xlab = "GDP (logged)", ylab = "Life Expectancy", zlab = "Happiness")
```
## Happiness by Year
```{r}
ggplot(data=happy)+  geom_point(aes(x=year, y=HPI, colour = region))
```
